---
title: "HW 7"
author: "Benson Wu"
date: "11/22/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(readxl)
library(survival)
library(KMsurv)
library(dplyr)
```


# Problem 8.12

```{r}
#Load data
data("std") 
```

```{r}
model1 <- coxph(Surv(time, rinfct)~  race + marital + age + yschool + 
                   npartner + os12m + rs12m + factor(condom) + abdpain + 
                   discharge + dysuria + itch + lesion + rash + lymph + 
                   factor(iinfct) , data = std, ties = "breslow")

#Use the step algorithm to find the most predictive factors 
step(model1)


```


```{r}

model1_step <- coxph(Surv(time, rinfct) ~ yschool + npartner + os12m + 
                      factor(condom) + abdpain + factor(iinfct), 
                      data = std, ties = "breslow")
summary(model1_step)
```



After the step algorithm, years of schooling, oral sex within 12 months, abdominal pain, and having chlamydia were significant predictors of time until reinfection. 


# Problem 9.3

## Part A
```{r}
time <-c(1, 63, 105, 129, 182, 216, 250, 262, 301, 301, 342, 354,
356, 358, 380, 383, 383, 388, 394, 408, 460, 489, 499, 523,
524, 535, 562, 569, 675, 676, 748, 778, 786, 797, 955, 968,
1000, 1245, 1271, 1420, 1551, 1694, 2363, 2754, 2950,
17, 42, 44, 48, 60, 72, 74, 95, 103, 108, 122, 144,
167, 170, 183, 185, 193, 195, 197, 208, 234, 235, 254, 307,
315, 401, 445, 464, 484, 528, 542, 547, 577, 580, 795, 855,
1366, 1577, 2060, 2412, 2486, 2796, 2802, 2934, 2988)
delta <- c(rep(1, 43), rep(0,2),rep(1, 39), rep(0,6))
treatment <- c(rep("chemo",45), rep("radio",45))

df <- as.data.frame(cbind(time, delta, treatment))

df$time <- as.numeric(df$time)
df$delta <- as.numeric(df$delta)

model1 <- coxph(Surv(time, delta)~factor(treatment), 
                  data = df, ties = "breslow")
summary(model1)


# 95% confidence interval
round(exp((model1$coef[1]+c(-1,1)*1.96*sqrt(model1$var[1,1]))),6)
```
We fail to reject the null that the survival is different between the two treatment regimes. 

## Part B

```{r}
df$treatment_indicator <- ifelse(df$treatment=="chemo",1,0)


event.times <- sort(unique(df$time[df$delta == 1]))

#restructure the data into a counting process-like dataset
df_restructured <- survSplit(data = df, cut = event.times, end = "time", start = 't0', event = "delta")

#Create time dependent covariate
df_restructured$tdc <- df_restructured$treatment_indicator*log(df_restructured$time)

model2<- coxph(Surv(t0, time, delta)~ treatment + tdc, 
                  data = df_restructured, ties = "breslow")
summary(model2)

#Testing b2 = 0
C  <- c(0, 1)
b0 <- c(0, 0)
b  <- model2$coefficients
V  <- model2$var
wald <- t(C %*% b - C %*% b0) %*% solve(t(C) %*% V %*% C) %*% (C %*% b - C %*% b0)
round(cbind(wald, 1-pchisq(wald, 1)),6)
```


## Part C

```{r}
#Pg. 306
loglik <- as.numeric(length(event.times))

#Warnings may occur due to X being singular. (that is a problem with this method)
for(i in 1:length(event.times)){
  df_restructured$tdc1 <- ifelse(df_restructured$time > event.times[i], df_restructured$treatment, 0)
  fit <- coxph(Surv(t0, time, delta) ~ treatment + tdc1, data = df_restructured, ties = 'breslow')
  loglik[i] <- fit$loglik[2]
}

cbind(event.times, loglik)
opt_tau <- event.times[which.max(loglik)]
opt_tau
```

## Part D

```{r}
df_restructured$Z1 <- ifelse(df_restructured$time > opt_tau, df_restructured$treatment_indicator, 0)
df_restructured$Z2 <- ifelse(df_restructured$time <= opt_tau, df_restructured$treatment_indicator, 0)

model3 <- coxph(Surv(t0, time, delta) ~  Z1 + Z2, data = df_restructured, ties = 'breslow')
summary(model3)
```
The RR of Z1 is 1.894 and the RR of Z2 is 0.241. In model A, the RR was 1.113 for those who received radiotherapy. 


# Question 9.5

## Part A
```{r}
data(larynx) 
larynx <- larynx %>% 
          mutate(diag_before75 = ifelse(diagyr<75, 1, 0))
model1 <- coxph(Surv(time, delta) ~  age + factor(stage) + strata(diag_before75),
                data = larynx, ties = 'breslow')
summary(model1)
```

## Part B

```{r}
after75 <- larynx %>% filter(diag_before75==0)
before75 <- larynx %>% filter(diag_before75==1)

fit0 <- coxph(Surv(time, delta) ~  age + factor(stage) + strata(diag_before75), data = after75, ties = 'breslow') 
fit1 <- coxph(Surv(time, delta) ~  age + factor(stage) + strata(diag_before75), data = before75, ties = 'breslow')
X2 <- -2*(model1$loglik[2] - (fit0$loglik[2] + fit1$loglik[2]))
X2
1 - pchisq(X2, df = 4)
```
Fail to reject the null.

## Part C
#Perform Wald Test
```{r}

C  <- rbind(c(1, 0, 0, 0, -1, 0, 0, 0),
                 c(0, 1, 0, 0, 0, -1, 0, 0),
                 c(0, 0, 1, 0, 0, 0, -1, 0),
                 c(0, 0, 0, 1, 0, 0, 0, -1))
b0 <- c(0, 0, 0, 0, 0, 0, 0, 0)
b <- c(fit0$coefficients, fit1$coefficients)
V <- rbind(cbind(fit0$var, matrix(0, nrow = nrow(fit1$var),
                                          ncol = ncol(fit1$var))),
                 cbind(matrix(0, nrow = nrow(fit1$var),
                              ncol = ncol(fit0$var)), fit1$var)
                 ) 
1 - pchisq(t(C %*% b- C %*% b0) %*% 
             solve(C %*% V%*% t(C)) %*% 
             (C %*% b- C %*% b0), 4)

```

We also fail to reject the null.



# Problem 9.8

## Part A
```{r}
data("burn")

cut.points <- sort(unique(burn$T3[burn$D3 == 1]))
burn_restructured <- survSplit(data = burn, cut = cut.points, end = "T3", start = 't0', event = "D3")

burn_restructured$tdw <- ifelse(burn_restructured$T3 >= burn_restructured$T1 & burn_restructured$D1 == 1, 1, 0)

model1 <- coxph(Surv(t0, T3, D3) ~ Z1 + Z2 + Z3 + tdw, data = burn_restructured, ties = 'breslow')
summary(model1)
```


## Part B
```{r}
#Create a TDC
burn_restructured$tdc_1 <- ifelse(burn_restructured$T3 >= burn_restructured$T2 & burn_restructured$D2 == 1, 1, 0)

model2 <- coxph(Surv(t0, T3, D3) ~ Z1 + Z2 + Z3 + tdc_1, data = burn_restructured, ties = 'breslow')
summary(model2)
```

## Part C
```{r}
model3 <- coxph(Surv(t0, T3, D3) ~ Z1 + Z2 + Z3 + Z4 + Z5 +Z6 + Z7 + Z8 + Z9 + Z10 + factor(Z11) + tdc_1 + tdw, data = burn_restructured, ties = 'breslow')

step(model3)
```

```{r}
#Final model 
model4 <- coxph(formula = Surv(t0, T3, D3) ~ Z1 + Z2 + Z3 + factor(Z11), 
    data = burn_restructured, ties = "breslow")
summary(model4)
```
## Part D
Make an inference about the time until infection among those individuals who had a prophylactic antibiotic treatment administered.
Adjust for all other explanatory covariates, as needed. Test for proportional hazards, and deal with any variables with nonproportional hazards, as you deem appropriate.
```{r}
burn_restructured_anti <- filter(burn_restructured, D2==1)
model5 <- coxph(Surv(t0, T3, D3) ~ Z1 + Z2 + Z3 + Z4 + Z5 + 
                  Z6 + Z7 + Z8 + Z9 + Z10 + factor(Z11) + 
                  tdc_1 + tdw, data = burn_restructured_anti, ties = 'breslow')
step(model5)

#Final model 
summary(coxph(Surv(t0, T3, D3) ~ Z5 + Z6 + Z9 + tdc_1, data = burn_restructured_anti, ties = 'breslow'))
```


# Question 11.1
```{r}
data("larynx") 
```

## Part C
```{r}
model1 <- coxph(Surv(time,delta) ~ factor(stage) + age, data = larynx, 
                method = 'breslow')


#Get Cox-Snell residual based on Martingale residuals
mg.residual <- resid(model1, type = "martingale")
cs.residual <- larynx$delta - mg.residual

#Graphical Plot
fit.cs <- survfit(Surv(cs.residual, larynx$delta) ~ 1) #Get Kaplan-Meier estiamtes
H.cs   <- cumsum(fit.cs$n.event/fit.cs$n.risk)
plot(fit.cs$time, H.cs, type='s', col='blue', 
     main = 'Cox-Snell Residual Plot', 
     xlab = 'Residual', ylab = 'Nelson-Aalen Cum. Hazard') 
#Note here that 'time' is the value of the Cox-Snell residual
abline(0, 1, col='red',  lty = 2)
```


# Question 11.3

## Part A
```{r}
data("larynx")

model1 <- basehaz(coxph(Surv(time,delta) ~ strata(factor(stage)) + age, data = larynx,  method = 'breslow'))
model1

plot(log(model1$hazard[model1$strata == 1])~ model1$time[model1$strata == 1], type = 's',
     ylab = 'Log Cumulative Hazard', xlab = 'Time', main = 'Log H(t) vs. Time',
     col = 'blue', lty = 1, xlim = c(0, 11), ylim = c(-5, 5))
lines(log(model1$hazard[model1$strata == 2]) ~  model1$time[model1$strata == 2],
      col = 'red', lty = 2, type = 's')
lines(log(model1$hazard[model1$strata == 3]) ~  model1$time[model1$strata == 3],
      col = 'green', lty = 3, type = 's')
lines(log(model1$hazard[model1$strata == 4]) ~  model1$time[model1$strata == 4],
      col = 'black', lty = 4, type = 's')
legend('bottomright', c('Stage 1', 'Stage 2', "Stage 3", "Stage 4"), col = c('blue', 'red', "green", "black"), lty = c(1,2,3,4), bty = 'n')
```


## Part B
```{r}
H1 <- model1$hazard[model1$strata == 1]
H2 <- model1$hazard[model1$strata == 2]
H3 <- model1$hazard[model1$strata == 3]
H4 <- model1$hazard[model1$strata == 4]

t1 <- model1$time[model1$strata == 1]
t2 <- model1$time[model1$strata == 2]
t3 <- model1$time[model1$strata == 3]
t4 <- model1$time[model1$strata == 4]

plot(log(H2) - log(H1[1:15]) ~ t2, main = 'Log H(t) differences', ylab = 'Log Cumulative Hazard', xlab = 'Time', type = 's', xlim = c(0, 10), ylim = c(-4, 4))
lines(log(H3) - log(H1[1:23]) ~ t3,
      col = 'blue', lty = 3, type = 's')
lines(log(H4) - H1[1:12] ~ t4,
      col = 'green', lty = 3, type = 's')
legend('bottomright', c('Stage 2 - Stage 1', "Stage 3 - Stage 1", "Stage 4 - Stage 1"), col = c('black', "blue", "green"), lty = c(1, 2,3), bty = 'n')
```


## Part C
```{r}
reptime <- function(l, t){
  x <- numeric(max(t))
  for(i in min(t):max(t)){
    diff <- i - t
    diff <- diff[diff >= 0]
    x[i] <- l[which.min(diff)]
  }
  return(x)
}

H1 <- model1$hazard[model1$strata == 1]
H2 <- model1$hazard[model1$strata == 2]
H3 <- model1$hazard[model1$strata == 3]
H4 <- model1$hazard[model1$strata == 4]
t1 <- model1$time[model1$strata == 1]
t2 <- model1$time[model1$strata == 2]
t3 <- model1$time[model1$strata == 3]
t4 <- model1$time[model1$strata == 4]

H1 <- reptime(H1, t1)
H2 <- reptime(H2, t2)
H3 <- reptime(H3, t3)
H4 <- reptime(H4, t4)


plot(H2[1:15] ~ H1[1:15], main = 'Anderson Plot', ylab = 'Cumulative Hazard', xlab = 'Cumulative Hazard', type = 's', xlim = c(0, 1), ylim = c(0, 1))
lines(H3[1:23] ~ H1[1:23],
      col = 'blue', lty = 3, type = 's')
lines(H4[1:12] ~ H1[1:12],
      col = 'green', lty = 3, type = 's')
abline(0, 1, col='red',  lty=2)
legend('bottomright', c('Stage 2', "Stage 3", "Stage 4"), col = c('black', "blue", "green"), lty = c(1, 2,3), bty = 'n')
```

